---
title: "BRSM Project"
author: "Team statStat: Prerak Srivastava, 
        Kishore Kumar, 
        Praneetha Gokul"
date: "27/3/2023"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

# Data Analysis

## Importing Libraries and Data Preprocessing

```{r}
# import libraries
library(readxl)
library(ggplot2)
library(car)
library(ppcor)
library(psych)
library(reshape2)
library(haven)
library(dplyr)
library(zoo)
```

```{r}
data1 <- read_sav("data/EEfRT Data for OSF_Long.sav")
data2 <- read_sav("data/EEfRT Data for OSF_Wide.sav")

data <- full_join(data2, data1, by = "ID", copy = FALSE)
summary(data)
```

```{r}
#removed between subject rows
new_data <- data %>% filter(ID == unique(data$ID))
```

```{r}
#created array without NA values
filtered_new_data <- na.omit(data$EEfRT_ProportionHardTasksSelected)

#removed subject ID with NA values for the task and lifetime substance use
data_update <- new_data[c(-42, -43, -44, -45, -46, -47), ]
data_update <- subset(data_update, !(is.na(LifetimeSubstanceDisorder)))

#assigned comorbid conditions as factored variables
data_update$ComorbidAnxietyDisorder <- as.factor(
    data_update$ComorbidAnxietyDisorder
)
data_update$LifetimeSubstanceDisorder <- as.factor(
    data_update$LifetimeSubstanceDisorder
)

#SCID was used to assess comorbidity
#There were only 2 factor levels for both comorbidities
#(1 = no presence, 3 = presence)
#These factor levels were assigned as (1 = 0, 3 = 1)
levels(data_update$ComorbidAnxietyDisorder)
levels(data_update$LifetimeSubstanceDisorder)
levels(data_update$ComorbidAnxietyDisorder) <- c(0, 1)
levels(data_update$LifetimeSubstanceDisorder) <- c(0, 1)
```

```{r}
#Removed extra columns
data_update <- subset(data_update, select = -c(7, 8, 39))

#Renamed columns
data_update <- data_update %>% rename(
    Age = Age.y,
    Gender = Gender.y
)

#converted to factored variable for analyses
data_update$Gender <- as.factor(data_update$Gender)
levels(data_update$Gender)
```

## Normality Tests

```{r}
shapiro.test(data_update$EEfRT_ProportionHardTasksSelected)
shapiro.test(data_update$Reward_Level)
shapiro.test(data_update$WASSUP_FinSuccess)
shapiro.test(data_update$WASSUP_PopFame)
```

Plotting the data to visualize the normality/abnormality
```{r}

ggplot(data_update, aes(x = EEfRT_ProportionHardTasksSelected)) +
  geom_histogram()
ggplot(data_update, aes(x = Reward_Level)) +
  geom_histogram()
ggplot(data_update, aes(x = WASSUP_FinSuccess)) +
  geom_histogram()
ggplot(data_update, aes(x = WASSUP_PopFame)) +
  geom_histogram()
```

```{r}
# filter all the data from data_update where comorbid anxiety disorder is 1
data_update_anxiety <- data_update[data_update$ComorbidAnxietyDisorder == 1, ]
data_update_substance <- data_update[data_update$LifetimeSubstanceDisorder == 1, ]
data_update_both <- data_update[data_update$ComorbidAnxietyDisorder == 1 &
    data_update$LifetimeSubstanceDisorder == 1, ]
data_update_neither <- data_update[data_update$ComorbidAnxietyDisorder == 0 &
    data_update$LifetimeSubstanceDisorder == 0, ]
```

```{r}
head(data_update_neither, 10)
head(data_update_anxiety, 10)
head(data_update_substance, 10)
head(data_update_both, 10)
```

## Exploratory Correlation Tests

```{r}
#Correlation tests
cor.test(
    data_update$EEfRT_ProportionHardTasksSelected, data_update$Reward_Level
)

cor.test(
    data_update$EEfRT_ProportionHardTasksSelected, data_update$LithiumDosage
)

cor.test(
    data_update$EEfRT_ProportionHardTasksSelected, data_update$WASSUP_FinSuccess
)

cor.test(
    data_update$EEfRT_ProportionHardTasksSelected, data_update$Age
)

cor.test(
    data_update$EEfRT_ProportionHardTasksSelected,
    as.numeric(data_update$Gender)
)

cor.test(
    data_update$WASSUP_FinSuccess, data_update$WASSUP_PopFame
)
```

## Correlation Heatmap

## Correlation Heatmap

```{r}
# Removed the ID column
data_update_1 <- data_update[, -1]

# Converted all the NAs to -1
data_update_1[is.na(data_update_1)] <- -1

# Converted each element in the data frame to numeric
data_update_1 <- as.data.frame(lapply(data_update_1, as.numeric))
data_update_1 <- na.aggregate(data_update_1, FUN = median)

cor_data <- cor(data_update_1)

melt_data_update <- melt(cor_data, varnames = c("Variable1", "Variable2"))
melt_data_update <- melt_data_update[order(-melt_data_update$value), ]
```

```{r}
# Ploting heatmap with sorted variables
ggplot(melt_data_update, aes(x = Variable1, y = Variable2)) +
  geom_tile(aes(fill = value), color = "white") +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
  midpoint = 0, na.value = "gray") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 5.5),
    axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5, size = 5.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14)
  ) +
  labs(
    x = "Variable",
    y = "Variable",
    fill = "Correlation"
  )
```

None of the correlation tests were significant. Correlation between EEfRT proportion and Reward Level r =
-0.056, p > .01 had a negative and weak relationship. Similarly, correlation between EEfRT proportion and
Lithium Dosage r = -0.13, p > .01 had a negative and weak relationship. Additionally, correlation between
EEfRT proportion and Age r = -0.137, p > .01 had a negative and weak relationship.

## Exploratory T-tests

```{r}
#Independent sample t-tests
t.test(EEfRT_ProportionHardTasksSelected ~ Gender, data_update)
t.test(Reward_Level ~ Gender, data_update)

#Independent sample t-tests of DVs with IV = Comorbid Anxiety Disorder
t.test(EEfRT_ProportionHardTasksSelected ~ ComorbidAnxietyDisorder, data_update)
t.test(WASSUP_PopFame ~ ComorbidAnxietyDisorder, data_update)
t.test(WASSUP_FinSuccess ~ ComorbidAnxietyDisorder, data_update)

#Independent sample t-tests of DVs with IV = Lifetime Substance use Disorder
t.test(
    EEfRT_ProportionHardTasksSelected ~ LifetimeSubstanceDisorder, data_update
)
t.test(WASSUP_PopFame ~ LifetimeSubstanceDisorder, data_update)
t.test(WASSUP_FinSuccess ~ LifetimeSubstanceDisorder, data_update)
```

```{r}
# create a data frame with the p-values of the t-tests
p_values <- data.frame(
  IV = c("Gender", "Gender", "Gender", "Gender",
        "Comorbid Anxiety", "Comorbid Anxiety",
        "Comorbid Anxiety", "Comorbid Anxiety", # nolint
        "Lifetime Substance Use", "Lifetime Substance Use",
        "Lifetime Substance Use", "Lifetime Substance Use",
        "0.05", "0.05", "0.05", "0.05"
    ), # nolint
  DV = c("EEfRT", "Reward_Level", "WASSUP Fin", "WASSUP Pop",
        "EEfRT", "WASSUP Pop", "WASSUP Fin", "Reward_Level", # nolint
         "EEfRT", "WASSUP Pop", "WASSUP Fin", "Reward_Level",
         "EEfRT", "WASSUP Pop", "WASSUP Fin", "Reward_Level"), # nolint
  p_value = c(t.test(EEfRT_ProportionHardTasksSelected ~ Gender, data_update)$p.value, # nolint
              t.test(Reward_Level ~ Gender, data_update)$p.value, # nolint
              t.test(WASSUP_FinSuccess ~ Gender, data_update)$p.value, # nolint
              t.test(WASSUP_PopFame ~ Gender, data_update)$p.value, # nolint
              t.test(EEfRT_ProportionHardTasksSelected ~ ComorbidAnxietyDisorder, data_update)$p.value, # nolint
              t.test(WASSUP_PopFame ~ ComorbidAnxietyDisorder, data_update)$p.value, # nolint
              t.test(WASSUP_FinSuccess ~ ComorbidAnxietyDisorder, data_update)$p.value, # nolint
              t.test(Reward_Level ~ ComorbidAnxietyDisorder, data_update)$p.value, # nolint
              t.test(EEfRT_ProportionHardTasksSelected ~ LifetimeSubstanceDisorder, data_update)$p.value, # nolint
              t.test(WASSUP_PopFame ~ LifetimeSubstanceDisorder, data_update)$p.value, # nolint
              t.test(WASSUP_FinSuccess ~ LifetimeSubstanceDisorder, data_update)$p.value, # nolint
              t.test(Reward_Level ~ LifetimeSubstanceDisorder, data_update)$p.value, # nolint
              0.05, 0.05, 0.05, 0.05) # nolint
)
```

```{r}
ggplot(
    p_values, aes(x = p_value, y = reorder(IV, p_value))) +
    geom_point(aes(color = p_value), size = 3, alpha = 0.8) +
    facet_wrap(~DV, nrow = 3, scales = "free_y") +
    scale_color_gradient(low = "magenta", high = "purple") +
    labs(
        title = "Support for Null Hypothesis",
        subtitle = "p-values for Independent Samples t-tests",
        x = "p-value",
        y = "Independent Variable"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 8, angle = 23, hjust = 1)) +
    geom_vline(xintercept = 0.05, color = "red", linetype = "dashed"
)
```

None of the t-tests were significant. A t-test between EEfRT and Gender did not exhibit
any significant difference between males (M = 0.582) and females (M = 0.476) on the EEfRT
proportion t(30) = 1.33, p = 0.193. A t-test between Reward Level and Gender did not exhibit any
significant difference between males (M =2.754) and females (M = 2.351) on the Reward level,
t(34) = 1.508, p = 0.14.

## ANOVA and Tukey HSD

```{r}
#Interaction between ComorbidAnxietyDisorder
#and LifetimeSubstanceDisorder on EEfRT_ProportionHardTasksS
model1 <- aov(
    EEfRT_ProportionHardTasksSelected ~
    ComorbidAnxietyDisorder * LifetimeSubstanceDisorder,
    data = data_update
)
summary(model1)
```

```{r}
TukeyHSD(model1)
```

```{r}
#Interaction between ComorbidAnxietyDisorder
# and LifetimeSubstanceDisorder on WASSUP_PopFame
model2 <- aov(
    WASSUP_PopFame ~ ComorbidAnxietyDisorder * LifetimeSubstanceDisorder,
    data = data_update
)
summary(model2)
```

```{r}
TukeyHSD(model2)
```

```{r}
#Interaction between ComorbidAnxietyDisorder
# and LifetimeSubstanceDisorder on WASSUP_FinSuccess
model3 <- aov(
    WASSUP_FinSuccess ~ ComorbidAnxietyDisorder * LifetimeSubstanceDisorder,
    data = data_update
)
summary(model3)
```

```{r}
TukeyHSD(model3)
```

## Linear Regression

```{r}
reg1.model <- lm(
    EEfRT_ProportionHardTasksSelected ~
    ComorbidAnxietyDisorder + WASSUP_FinSuccess,
    data = data_update
)

summary(reg1.model)
vif(reg1.model)
ncvTest(reg1.model)
plot(reg1.model)
AIC(reg1.model)
```

```{r}
reg3.model <- lm(
    EEfRT_ProportionHardTasksSelected ~ WASSUP_FinSuccess + ComorbidAnxietyDisorder,
    data = data_update
)

summary(reg3.model)
vif(reg3.model)
ncvTest(reg3.model)
plot(reg3.model)
AIC(reg3.model)
```

```{r}
# BEST FIT MODEL
reg.model4 <- lm(
    EEfRT_ProportionHardTasksSelected ~ LifetimeSubstanceDisorder +
    WASSUP_FinSuccess,
    data = data_update_anxiety
)

summary(reg.model4)
vif(reg.model4)
ncvTest(reg.model4)
plot(reg.model4)
AIC(reg.model4)
```

```{r}
reg.model15 <- lm(
    EEfRT_ProportionHardTasksSelected ~ LifetimeSubstanceDisorder +
    WASSUP_FinSuccess,
    data = data_update
)

summary(reg.model15)
vif(reg.model15)
ncvTest(reg.model15)
plot(reg.model15)
AIC(reg.model15)
```

```{r}
reg.model8 <- lm(
    EEfRT_ProportionHardTasksSelected ~ LifetimeSubstanceDisorder + LithiumDosage +
    WASSUP_FinSuccess,
    data = data_update
)

summary(reg.model8)
vif(reg.model8)
ncvTest(reg.model8)
plot(reg.model8)
AIC(reg.model8)
```

## Best Fit Line (Replication of Paper)

```{r}
# plot the regression line between EEfRT and WASSUP_FinSuccess with the data points
ggplot(data_update, aes(x = WASSUP_FinSuccess, y = EEfRT_ProportionHardTasksSelected)) +
    geom_point() +
    geom_smooth(method = "lm")
```

```{r}
# plot the regression line between EEfRT and WASSUP_PopFame with the data points
ggplot(data_update, aes(x = WASSUP_PopFame, y = EEfRT_ProportionHardTasksSelected)) +
    geom_point() +
    geom_smooth(method = "lm")
```

## Plots

### Violin Plots

```{r}
# Boxplot
# in the fill, replace 0 with No and yes with 1
ggplot(
        data_update,
        aes(
            x = ComorbidAnxietyDisorder,
            y = EEfRT_ProportionHardTasksSelected,
            fill = factor(LifetimeSubstanceDisorder)
        )
    ) +
    geom_violin() +
    theme_minimal() +
    labs(
        x = "Comorbid Anxiety Disorder",
        y = "EEfRT Proportion Hard Tasks Selected",
        fill = "Lifetime Substance Use Disorder"
    ) +
    scale_fill_discrete(
        name = "Lifetime Substance Use Disorder",
        labels = c("No Substance Use Disorder", "Substance Use Disorder")
    )
```

```{r}
ggplot(
    data_update,
    aes(
        x = ComorbidAnxietyDisorder,
        y = EEfRT_ProportionHardTasksSelected,
        fill = ComorbidAnxietyDisorder
    )
) +
    geom_violin()+
    theme_bw() +
    theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "none") +
    scale_fill_brewer(palette = "Paired") +
    labs(title = "Comorbid Anxiety Disorder Proportion of
    Hard Tasks Selected on EEfRT",
    x = "Comorbid Anxiety Disorder",
    y = "EEfRT proportion"
)
```

```{r}
ggplot(
        data_update,
        aes(
            x = ComorbidAnxietyDisorder,
            y = WASSUP_PopFame,
            fill = ComorbidAnxietyDisorder
        )
    ) +
    geom_violin() +
    theme_bw() +
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none"
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(
        title = "Comorbid Anxiety Disorder WASSUP Popularity Fame",
        x = "Comorbid Anxiety Disorder",
        y = "WASSUP Popularity Fame"
    )
```

```{r}
ggplot(
        data_update,
        aes(
            x = ComorbidAnxietyDisorder,
            y = WASSUP_FinSuccess,
            fill = ComorbidAnxietyDisorder
        )
    ) +
    geom_violin() +
    theme_bw() +
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none"
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(
        title = "Comorbid Anxiety Disorder WASSUP Financial Success",
        x = "Comorbid Anxiety Disorder",
        y = "WASSUP Financial Success"
    )
```

```{r}
ggplot(
        data_update,
        aes(
            x = LifetimeSubstanceDisorder,
            y = EEfRT_ProportionHardTasksSelected,
            fill = LifetimeSubstanceDisorder
        )
    ) +
    geom_violin() +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)
    ) +
    labs(
        x = "Lifetime Substance Use Disorder",
        y = "EEfRT Proportion Hard Tasks Selected",
        fill = "Lifetime Substance Use Disorder"
    )
```

```{r}
# plot between lifetime substance and fame
ggplot(
        data_update,
        aes(
            x = LifetimeSubstanceDisorder,
            y = WASSUP_PopFame,
            fill = LifetimeSubstanceDisorder
        )
    ) +
    geom_violin() +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)
    ) +
    labs(
        x = "Lifetime Substance Use Disorder",
        y = "WASSUP Popularity Fame",
        fill = "Lifetime Substance Use Disorder"
    )
```

```{r}
# plot between life substance and financial success
ggplot(
        data_update,
        aes(
            x = LifetimeSubstanceDisorder,
            y = WASSUP_FinSuccess,
            fill = LifetimeSubstanceDisorder
        )
    ) +
    geom_violin() +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)
    ) +
    labs(
        x = "Lifetime Substance Use Disorder",
        y = "WASSUP Financial Success",
        fill = "Lifetime Substance Use Disorder"
    )
```

### Interaction Plots

```{r}
ggplot(
    data_update,
    aes(
        x = ComorbidAnxietyDisorder,
        y = EEfRT_ProportionHardTasksSelected,
        color = factor(LifetimeSubstanceDisorder)
    )
) +
  stat_summary(
    fun = mean,
    geom = "point", size = 3, shape = 21, fill = "white") +
    stat_summary(
        fun = mean,
        geom="line",
        size = 0.5,
        aes(group = factor(LifetimeSubstanceDisorder))
    ) +
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", width = 0.2
    ) +
    labs(
        x = "Comorbid Anxiety",
        y = "EEfRT proportion",
        color = "Lifetime Substance Use Disorder"
    ) +
    scale_color_manual(
        values = c("red", "green"),
        name = "Lifetime Substance Use Disorder",
        labels = c("No", "Yes")
    ) +
    ggtitle("Interaction effect between Comorbid Anxiety and
            Lifetime Substance Use on EEfRT task proportion") +
    theme_minimal() +
    theme(plot.title = element_text(size=14, face="bold")) +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
    )
```

```{r}
# Plot interaction effect of between
# Comorbid Anxiety and Lifetime Substance Use on WASSUP Popular Fame Scale
ggplot(
    data_update,
    aes(x = ComorbidAnxietyDisorder, y = WASSUP_PopFame,
    color = factor(LifetimeSubstanceDisorder))) +
    stat_summary(fun = mean, geom = "point", size = 3,
    shape = 21, fill = "white") +
  stat_summary(fun = mean, geom="line", size = 0.5,
  aes(group = factor(LifetimeSubstanceDisorder))) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
  width = 0.2) +
  labs(x = "Comorbid Anxiety", y = "WASSUP_PopFame",
  color = "Lifetime Substance Use Disorder") +
  ggtitle("Interaction effect between Comorbid Anxiety
  and Lifetime Substance Use on WASSUP PopFame Scale") +
  scale_color_manual(values = c("red", "green"),
  name = "Lifetime Substance Use Disorder", labels = c("No", "Yes")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14,
  face = "bold")) +
  theme(legend.position = "bottom",
  legend.title = element_text(size = 10),
  legend.text = element_text(size = 8))
```

```{r}
ggplot(data_update, aes(x = ComorbidAnxietyDisorder,
y = WASSUP_FinSuccess, color = factor(LifetimeSubstanceDisorder))) +
  stat_summary(fun = mean, geom = "point", size = 3,
  shape = 21, fill = "white") +
  stat_summary(fun=mean, geom="line", size=0.5,
  aes(group = factor(LifetimeSubstanceDisorder))) +
  stat_summary(fun.data = mean_cl_normal,
  geom = "errorbar", width = 0.2) +
  labs(x = "Comorbid Anxiety", y = "WASSUP_FinSuccess",
  color = "Lifetime Substance Use Disorder") +
  scale_color_manual(values = c("red", "green"),
  name = "Lifetime Substance Use Disorder", labels = c("No", "Yes")) +
  ggtitle("Interaction effect between Comorbid Anxiety
  and Lifetime Substance Use on WASSUP FinSuccess Scale") +
  theme_minimal() +
  theme(plot.title = element_text(size=14, face="bold")) +
  theme(legend.position = "bottom",
  legend.title = element_text(size = 10), legend.text = element_text(size = 8))
```